---
title: "P8105_hw2_ys3394"
author: "Yifan Su"
date: "9/30/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

#### Read and clean the Mr. Trach Wheel dataset.

```{r trash_df, message = FALSE, collapse = TRUE}
trash_df =
  read_xlsx( # read the xlsx.
    "./data2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% # drop NA only in the dumpster column
  mutate(
    sport_balls = round(sports_balls),
    sport_balls = as.integer(sports_balls)
  ) %>% 
  relocate(dumpster, year, month, everything())

head(trash_df)
```

#### Read and clean precipitation data for 2017 and 2018.

```{r pre_2017&2018, message = FALSE}
# Load pre_2018 dataset
pre_2018 = 
  read_xlsx(
    "./data2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1) %>%  # skip the first row of dataset
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>%  # add a column showing the year
  relocate(year)

# Load pre_2017 dataset
pre_2017 = 
  read_xlsx(
    "./data2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Then, combine two precipitation datasets.

```{r pre_df, message = FALSE, collapse = TRUE}
# Create a month tibble, and left_join original pre_df data
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name)

pre_df =
  bind_rows(pre_2018, pre_2017) %>%
  left_join(month_df, by = "month")

str(pre_df)
```

```{r include = FALSE}
trash_dfm = 
  filter(trash_df, year == "2017")
```

Mr.Trash Wheel is a vessel removes trash in Baltimore, Maryland since 2014. The dataset is about the information stored in the dumpster, including key variables **_`r names(trash_df)`_**. For the data of precipitation in 2018, there are `r nrow(pre_2018)` rows, and the total precipitation in 2018 is `r sum(pull(pre_2018, total))`. The median number of sports balls in the dumpster in 2017 is `r median(pull(trash_dfm, sports_balls))`.


## Problem 2

#### Read and clean the NYC Transit dataset.

```{r NYC_df, message = FALSE, collapse = TRUE}
# Remove previous variables from the environment
rm(list = ls()) 

NYC_df = 
  read_csv(
    "./data2/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%   
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>%  # select those columns
  mutate(entry = if_else(entry == "YES", 1, 0))

head(NYC_df)
```

In the NYC transit dataset, there are key variables **_`r names(NYC_df)`_**. By now, I've read the csv file with the readxl package, select required columns, and converted the entry variable from the character variable to the logical variable. It can bee seen from the tibble bellow, which is head rows of `NYC_df`. The dataset is a `r NYC_df %>% nrow()` X `r NYC_df %>% ncol()` tibble. The dataset is not yet tidy, there are variables such as those inconsistent route numbers and names, can be modified.

* There are `r nrow(distinct(NYC_df, line, station_name))` distinct stations in the dataset.

* There `r nrow(filter(NYC_df, ada == "TRUE"))` stations that are ADA compliant. 

* The proportion of station entrances and exits without vending is `r NYC_df %>% filter(vending == "NO", entry == "1") %>% nrow / NYC_df %>% filter(vending == "NO") %>% nrow`.

#### Reformate data, make route number and route name distinct variables.

```{r NYC_tidy, message = FALSE, collapse = TRUE}
# Tidy the NYC_df to be NYC_tidy
NYC_tidy = 
  NYC_df %>% 
  mutate(  # convert route8:rout11 to character variables
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  ) %>% 
  pivot_longer(  # pivot the data longer
    route1:route11,
    names_to = "route_name",
    values_to = "route_num"
  ) %>% 
  drop_na(route_num)

head(NYC_tidy)
```

* There are `r NYC_tidy %>% filter(route_num == "A") %>% distinct(station_name, line) %>% count()` distinct stations that serve the A train. 

* Among thee stations, there are `r NYC_tidy %>% filter(route_num == "A") %>% distinct(station_name, line, ada) %>% filter(ada == "TRUE") %>% count()` stations that are ADA compliant.


## Problem 3

#### Clean three datasets

Clean the data in pols_month.csv.

```{r pols_df, warning = FALSE, message = FALSE, collapse = TRUE}
rm(list = ls())

month_df =
  tibble(
    month = 1:12,
    month_name = month.name)

# Tidy the pols_dy data
pols_df = 
  read_csv(
    "./data2/pols-month.csv") %>%   
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), "-") %>% 
  mutate(month = as.integer(month)) %>%  
  left_join(month_df, by = "month") %>%  # left_join the month tibble in the same way
  select(-month, -day) %>% 
  pivot_longer(
    cols = starts_with("prez_"),
    names_to = "president",
    values_to = "president_val",
    names_prefix = "prez_"
  ) %>% 
  relocate(year, month_name, everything()) # everythng(), commonly used

str(pols_df)
```

Clean the data in snp.csv using the similar process.

```{r snp_df, warning = FALSE, message = FALSE, collapse = TRUE}
# Tidy the snp_df data
snp_df = 
  read_csv(
    "./data2/snp.csv") %>%   
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), "/") %>% 
  mutate(month = as.integer(month)) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month) %>% 
  relocate(year, month_name, everything())

str(snp_df)
```

Tidy the unemployment.csv dataset.

```{r unemployment_df, warning = FALSE, message = FALSE, collapse = TRUE}
# Create another month tibble
month_df2 =
  tibble(
    month = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),
    month_name = month.name
  )
# Tidy the unemployment_df data
unemployment_df = 
  read_csv(
    "./data2/unemployment.csv") %>%   
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>% 
  mutate(year = as.character(year)) %>% 
  left_join(month_df2, by = "month") %>% 
  select(-month) %>% 
  relocate(year, month_name, everything())

str(unemployment_df)
```

#### Join three datasets by year and month

```{r comb_df, message = FALSE, collapse = TRUE}
comb_df =  # left_join three dfs
  left_join(pols_df, snp_df, by = c("year", "month_name")) %>% 
  left_join(unemployment_df, by = c("year", "month_name")) %>% 
 relocate(year, month_name, everything())

head(comb_df)
```

**Summary**

* In dataset pols_df, there is a `r nrow(pols_df)` X `r ncol(pols_df)` tibble, with the year range from `r min(pull(pols_df, year))` to `r max(pull(pols_df, year))`. Key variables in this dataframe are **_president, sen_dem_**, etc.

* In dataset snp_df, it has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. The year range is from year `r min(pull(snp_df, year))` to year `r max(pull(snp_df, year))`. Key variables of it consist of **_`r names(snp_df)`_** .

* In dataset unemployment_df, it's a `r nrow(unemployment_df)` X `r ncol(unemployment_df)` tibble with the range of year from `r min(pull(unemployment_df, year))` to year `r min(pull(unemployment_df, year))`. Key variables are **_`r names(unemployment_df)`_** .

* In the combined dataset, three tibbles that sharing variable `year` and `montn_name` are combine together by `left_join`. there are some key variables **_`r names(comb_df)`_** . The combined dataset has `r nrow(comb_df)` rows with `r ncol(comb_df)` columns. The range of the dataset is `r distinct(comb_df, year) %>% count()-1` years, from `r min(pull(comb_df, year))` to `r max(pull(comb_df, year))`.




